Synthetic data (confidentiality) checks

### 2020-12-04 mice_synthesizing/data_checks.Rmd

```{r, results = F, message = F, warning=F}
library(mice) # imputations
library(tidyverse) # (d)plyr
library(magrittr) # pipes
```

@@

```{r}
meth <- make.method(mice::boys)
meth['bmi'] <- "~I(wgt / (hgt/100)^2)"
pred <- make.predictorMatrix(boys)
pred[c("wgt", "hgt"), "bmi"] <- 0

data <- mice::boys %>% 
  mice(m = 1, method = meth, predictorMatrix = pred, print = FALSE, seed = 123) %>% 
  complete()
```
Now we have a complete dataset, we can overimpute the complete dataset, so that we have synthetic data. First, we set the synthesizing method to `cart`. Then, we use mice to generate 100 synthetic versions of the completed dataset.

```{r, cache = thomcache}
cart <- meth; cart[!names(meth) %in% 'bmi'] <- "cart"

syns <- data %>% 
  mice(m = 100, 
       method = cart, 
       predictorMatrix = pred, 
       where = matrix(TRUE, nrow(data), ncol(data)),
       print = FALSE,
       seed = 123)
```

Now we have 100 synthetic versions of the same dataset, we can create 100 completed datasets.

```{r}
completed <- complete(syns, action = "all")
```

Now we have 100 complete synthetic datasets, we can inspect the number of duplicate cases in every synthetic dataset with regard to the original dataset.

```{r}
duplicate_obs <- map(completed, function(x) bind_rows(data, x) %>% duplicated)

@@

It appears that in all 100 synthetic versions of the data, there is not a single observation that is an exact copy of the actually observed observations. Now, let's find out if the synthetic data resembles the actually observed data (univariately). That is, we compare the distributions of the variables in the observed and the synthetic data.

@@

>>>>

## Load packages

```{r, results = F, message = F, warning=F}
library(mice) # imputations
library(tidyverse) # (d)plyr
library(magrittr) # pipes
library(furrr) # parallel imputation
```

@@

# Create fully observed data
```{r}
nsyn <- 1000 # number of synthetic datasets

# mice uses the actually observed data for imputing bmi, which is why it is the same in every dataset, and about equal to the true value
meth <- make.method(mice::boys)
meth['bmi'] <- "~I(wgt / (hgt/100)^2)"
pred <- make.predictorMatrix(boys)
pred[c("wgt", "hgt"), "bmi"] <- 0

data <- mice::boys %>% 
  mice(m = 1, method = meth, predictorMatrix = pred, print = FALSE, seed = 123) %>% 
  complete()
```

Now we have a complete dataset, we can overimpute the complete dataset, so that we have synthetic data. First, we set the synthesizing method to `cart`. Then, we use mice to generate 100 synthetic versions of the completed dataset.

# Create `r nsyn` synthetic data sets
```{r, cache = thomcache}
cart <- meth
cart[!names(meth) %in% 'bmi'] <- "cart"

plan(multisession)
syns <- future_map(1:nsyn, function(x) {
  data %>%
    mice(m = 1,
         method = cart,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(data), ncol(data)),
         print = FALSE)
}, .progress = FALSE, .options = future_options(seed = as.integer(123)))

```

## Complete all synthetic datasets

Now we have `r nsyn` synthetic versions of the same dataset, we can create `r nsyn` completed datasets.

```{r}
completed <- map(syns, function(x) complete(x))
```

Now we have `r nsyn` complete synthetic datasets, we can inspect the number of duplicate cases in every synthetic dataset with regard to the original dataset.

## Exact duplicate rows

I excluded the variable `bmi`, because something's gone wrong with imputing it.
```{r}
duplicate_obs <- map(completed, function(x) {
  bind_rows(data %>% select(-bmi), x %>% select(-bmi)) %>% duplicated
  }
)

@@

It appears that in all `r nsyn` synthetic versions of the data, there are seven observations that are an exact copy of actually sampled observations.

However, if we look at the proportion of the observations that has gotten his own value replaced by a value within $0.02$ times the range of the data, we find that between $0.39$ and $0.56$ of the values gets a replacement that is highly similar (`bmi` not counted, as for `bmi` every observation got his own value imputed due to some settings).

```{r}
out <- map(completed, function(x) {
  syn_num_vars <- 
    x %>% 
      select(where(is.numeric))

  true_num_vars <- 
    data %>%
      select(where(is.numeric))

  map2(syn_num_vars, true_num_vars, function(z, y) {
    abs(z - y) < (max(y) - min(y)) / 50
  }) %>% bind_cols
})

map_dfr(out, function(x) {
  map(x, function(z) sum(z))
}) %>% colMeans() / nrow(data)
```

Then, we check how often the nearly exact same value is imputed.

```{r}
out <- map(completed, function(x) {
  syn_num_vars <- 
    x %>% 
      select(where(is.numeric))

  true_num_vars <- 
    data %>%
      select(where(is.numeric))

  map2(syn_num_vars, true_num_vars, function(z, y) {
    abs(z - y) < (max(y) - min(y)) / 10000
  }) %>% bind_cols
})

map_dfr(out, function(x) {
  map(x, function(z) sum(z))
}) %>% colMeans() / nrow(data)
```

```{r}
out_cat <- map(completed, function(x) {
  syn_cat_vars <- 
    x %>%
      select(-where(is.numeric))

  true_cat_vars <- 
    data %>%
      select(-where(is.numeric))

  map2(syn_cat_vars, true_cat_vars, function(z, y) {
    z == y
  }) %>% bind_cols
})

map_dfr(out_cat, function(x) {
  map(x, function(z) sum(z))
}) %>% colMeans() / nrow(data)
```

For the categorical data, the same holds. The majority of all imputations gets his own value replaced, rather than a new value.

## Plot imputed data against the observed data

Now, let's find out if the synthetic data resembles the actually observed data (univariately). That is, we compare the distributions of the variables in the observed and the synthetic data.

@@

<!--
# Adjusted settings

```{r, cache = thomcache, eval = F}
syns_adj <- future_map(1:nsyn, function(x) {
  data %>%
    mice(m = 1,
         maxit = 5,
         method = cart,
         minbucket = 3,
         cp = 1e-08,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(data), ncol(data)),
         print = FALSE)
}, .progress = FALSE, .options = future_options(seed = as.integer(123)))
```

```{r, eval = F}
completed_adj <- map(syns_adj, function(x) complete(x))

completed_adj %>%
  map(function(x) bind_rows(data, x) %>% duplicated) %>%
  map_dbl(sum) %>% 
  sum # total number of duplicated rows over all 100 synthetic datasets

completed_adj %>%
  map(function(x) bind_rows(data, x) %>% duplicated) %>%
  lapply(function(x) which(x)) %>%
  unlist() - nrow(data)
```

```{r, dpi = 300, fig.fullwidth = TRUE, message = F, eval = F}
completed_adj %>%
  map_dfr(function(x) x) %>%
  bind_rows(., data, .id = "Imputed") %>%
  mutate(Imputed = as.numeric(Imputed)) %>%
  select_if(is.numeric) %>%
  pivot_longer(cols = -"Imputed",
               names_to = c("Variable")) %>%
  mutate(Imputed = recode(Imputed, `1` = "Synthetic data", `2` = "Observed data")) %>%
  ggplot(., mapping = aes(x = value, fill = Imputed)) +
  geom_histogram(aes(y = ..density..), position = "dodge2", ) +
  facet_wrap(~ Variable, nrow = 3, ncol = 2, scales = "free") +
  theme_classic() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r, dpi = 300, fig.fullwidth = TRUE, message = F, eval = F}
completed_adj %>%
  map_dfr(function(x) x) %>%
  bind_rows(., data, .id = "Imputed") %>%
  select_if(negate(is.numeric)) %>%
  map_dfr(function(x) as.character(x)) %>%
  pivot_longer(cols = -"Imputed",
               names_to = c("Variable")) %>%
  mutate(Imputed = recode(Imputed, `1` = "Synthetic data", `2` = "Observed data")) %>%
  ggplot(., mapping = aes(x = value, fill = Imputed, group = Imputed)) +
  geom_bar(aes(y = ..prop..), position = "dodge2") +
  facet_wrap(~ Variable, nrow = 3, ncol = 2, scales = "free") +
  theme_classic() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

-->
